Very simplified likelihood

### 2023-12-01 inst/very_simplified_likelihood.Rmd
more work on probs

@@

Placing after $r$ of the $n$ clones of strength $b$ finish:

@@

where $B=1/b$.  In R:

```{r}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     out <- log(B-1) - sum(log(B + B+(n-r-1):(n-1)))
   } else {
     out <- (B-1)/prod(B+(n-r-1):(n-1))
   }
   return(out)
}

>>>>

Probability of the focal competitor coming first:

\[
\frac{a}{a+nb}=\frac{a}{n - a(n-1)a}=\frac{1-b}{1+(n-1)b}
\]

Coming second:
\[
\frac{b}{a+nb}\cdot\frac{a}{a+(n-1)b}
\]

Third:

\[
\frac{b}{a+nb}\cdot\frac{b}{a+(n-1)b}\cdot\frac{a}{a+(n-2)b}
\]

@@

Placing after $r<n$ of the $n$ clones of strength $b$ finish:

@@

where $B=1/b$.

## Placing last

Placing last is finishing  after $n$ clones finish:

\[
n=1\longrightarrow\mathcal{L}=b\]

\[
n=2\longrightarrow\mathcal{L}=\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^2}{1+b}
\]

\[
n=3\longrightarrow\mathcal{L}=\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^3}{(1+b)(1+2b)}
\]

\[
n=4\longrightarrow\mathcal{L}=\frac{b}{a+4b}\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=
\frac{b^4}{(1+b)(1+2b)(1+3b)}
\]

For $n=4$ we have equivalently

\[
\frac{1}{B(B+1)(B+2)(B+3)}=\frac{(B-1)!}{(B+3)!}=\frac{\Gamma(B)}{\Gamma(B+4)}
\]

\[
\frac{B-1}{(B+n-1)(B+n-2)\cdots(B+1)\cdot B\cdot(B-1)}
\]

(denominator has $n+1$ terms).  In R:

```{r}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){
       out <- lgamma(B)-lgamma(B+n)
     } else {
       out <- log(B-1) - sum(log(B + B+(n-r-1):(n-1)))
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}x
